Discrete Quantum Theories and Computing

### 2018-11-09 dissertation.tex
dissertation.lyx Updated according to Prof. Hanson's suggestion...

%% LyX 2.3.0 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,twoside]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{mmap}
\usepackage{comment}
\usepackage{framed}
\usepackage{mathrsfs}
\usepackage{setspace}
\usepackage{dsfont}
\usepackage{xeCJK}
\setCJKmainfont{新細明體}
\usepackage{unicode-math}

@@

\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{prop}{Proposition}[chapter]
\newtheorem{thm}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem*{method}{Method}

@@

\newcommand{\muB}{\ensuremath{\mu^{\mathrm{B}}}}
\newcommand{\muC}{\ensuremath{\mu^{\mathrm{C}}}}
\newcommand{\muF}{\ensuremath{\mu^{\mathrm{F}}}}
\newcommand{\barmuD}{\ensuremath{\bar{\mu}^{\mathrm{D}}}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\eventsC}{\ensuremath{\events_{C}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\mathbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\bf #1}}
\newcommand{\mathReal}{\mathbb{R}}
\newcommand{\mathComplex}{\mathbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\mathbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\mathrm{e}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\CP}[1]{\mathComplex\mathbf{P}^{#1}}  % Complex projective space
\newcommand{\DCP}[1]{\mathbf{D}\mathComplex\mathbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\Psibar}{\overline{\Psi}}
\newcommand{\cardp}[2]{#1 \, \slash\!\! \slash \, #2}
\def\round{\mathop{\rm round}\nolimits}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\def\C{{\mathbb{C}}}
\newcommand{\ultramodular}{\mathcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}

@@

@@

\usepackage{polyglossia}
\setdefaultlanguage[variant=american]{english}
\setotherlanguage{arabic}
\usepackage[style=numeric-comp,sorting=none,maxbibnames=10,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{prop.bib}
\begin{document}
%data for Title Page

\title{Discrete Quantum Theories and Computing}

\author{Yu-Tsung Tai\\
(戴淯琮)}

\date{\today}

\maketitle
%data for Acceptance Page
\committeeMember{Amr Sabry (\textarabic{عمرو صبري}), PhD} \committeeMember{Gerardo
Ortiz, PhD} \committeeMember{Dylan Paul Thurston, PhD} \committeeMember{Andrew
J. Hanson, PhD} \committeeMember{Shouhong Wang (汪守宏), PhD} \defensedate{Defense
Date} \acceptancepage

%data for Copyright Page
\cryear{2018}
\copyrightpage

\begin{dedication}I would like to dedicate my thesis to my parents,
Cheng-Tien Tai (戴振沺) and Feng-Ming Chang (張鳳鳴). Thanks for encouraging
me to study abroad and all your support during my Ph.D. study.\end{dedication}

\begin{acknowledgments}I want to thank Prof.~Gerardo Ortiz, Prof.~Amr
Sabry (\textarabic{عمرو صبري}), and Prof.~Andrew Hanson for publishing
three papers \cite{geometry2013,DQT2014,THOS2018} with me, and enormous
other things. These papers become the main part of my thesis, especially
most of Chapter~\ref{chap:DQT=000026DQC} is based on “Geometry
of discrete quantum computing”~\cite{geometry2013} and “Discrete
quantum theories” (DQT)~\cite{DQT2014}; most of Chapter~\ref{chap:QIVPM}
is based on “Quantum Interval-Valued Probability: Contextuality
and the Born Rule” (QIVPM) \cite{THOS2018}. Together with Prof.~Dylan
Thurston and Prof.~Shouhong Wang (汪守宏), I want to thank you for getting
together to understand what I did and passing my dissertation proposal
Spring 2017. I also thank Prof. Dylan Thurston for serving in my advisory
committee and my Tier 3 committee, with other advisory committee members
passing my Computer Science qualifying exam Spring 2016. Especially,
my survey of real computation in the qualifying exam inspired me the
comparison of quantum computer and analog computer in Chapter~\ref{chap:Introduction}.
Chapter~\ref{chap:Introduction} also contains the motivation in
my application for Rethinking Foundations of Physics 2017 workshop
reviewed by Prof. Amr Sabry (\textarabic{عمرو صبري}). I want to thank
Prof. Lawrence Moss for inviting me to present our results in the
interdisciplinary logic seminar and theory seminar in Computer Science,
co-chairing my Tier 3 committee, writing an assessment letter for
me, and drawing my attention on Prof.~Abramsky's paper~\cite{Abramsky2012},
which inspired me to merge quantum probability with discrete quantum
theories. And I thank John Gardiner for inspiring discussion on the
difficulties merging them~\cite{Gardiner2014}. These results are
improved and motivated the transition from DQT to QIVPM in Sec.~\ref{sec:Toward-IVPM}.
I also thank Prof.~Tom Lewis for helping me organize the unpublished
deterministic quantum algorithms in the term paper for SLST-T501 Academic
Writing course as presented in Sec.~\ref{discretequantumtheoryI}.
I want to thank Traci Nagle for going through the literature review
about finite precision measurement and contextuality discussed for
CSCI-Y790 Writing and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.
I also thank Elizabeth “Betsy” Merceron and Kexin “Casey”
Chen (谌可心) for teaching me English in general so that others can understand
me more easily.

\todo{Type a table for acknowledgment so that I would have to write
“I thank” so many times here?}Many people in both department
of Mathematics and Computer Science need to do something more to make
my double-major works smoothly. I need to thank all of them, especially
directors of graduate study in Mathematics, especially Prof. Christopher
Judge and Prof. Matthias Weber; directors of Ph.D. study in Computer
Science, especially Prof. Yuqing Melanie Wu (吴愈青) and Prof. Funda
Ergun.

@@

Finally, thank Jin-Ru Yang (楊謹如) for everything could not be written
here.\end{acknowledgments}

\begin{preface}
\begin{itemize}
\item The latest version of this thesis is in GitHub:
\[
\hspace{-4em}\textrm{https://github.com/yuttai/Contextuality-and-Nonlocality-in-Discrete-Quantum-Theory/blob/master/dissertation.pdf}
\]
\item Its source code is typeset using \LyX ~\cite{LyX} in:
\[
\hspace{-4em}\textrm{https://github.com/yuttai/Contextuality-and-Nonlocality-in-Discrete-Quantum-Theory/blob/master/dissertation.lyx}
\]
\item Any comments can be left in the Issue part of the GitHub Repository:
\[
\hspace{-4em}\textrm{https://github.com/yuttai/Contextuality-and-Nonlocality-in-Discrete-Quantum-Theory/issues}
\]
\end{itemize}
\end{preface}
\begin{abstract}
Our primary research interest is to build a quantum computing model
characterizing realistic quantum computers. While most of the quantum
computing models based on uncomputable numbers, that is, the continuum
of real numbers, most of the classical computers in our daily life
are digital instead of analog computers. This highlight the necessity
to investigate discrete models for quantum theory and computing. Specifically,
we start from replacing the continuum of complex numbers by the discrete
finite fields. Although we have fruitful results on their geometric
implications and computing powers, their probability models are still
not completely satisfactory. To address this issue, we further exploited
quantum interval-valued probability, and proved an imprecise version
of foundational results such as the Gleason and Kochen-Specker theorems.
\end{abstract}
\tableofcontents{}

@@

When we check the compatibility between quantum mechanics and computer
science, we found their fundamental assumptions are different. On
one hand, the quantum theory is based on infinitely precise real and
complex numbers. Its prediction could fit the physical reality by
utilizing error analysis technique although these prediction cannot
be completely faithful because of the measurement precision. On the
other hand, current computers mostly perform digital computation,
except some analog chips communicating with physical world~\cite{Camenzind2005}.
While the computability of digital parts is faithfully characterized
by its theoretical model, the Turing machine~\cite{Turing_1937},
the computability of analog chips in reality is far weaker than the
theoretical prediction of real computation \cite{Siegelmann1998,Ziegler2007,weihrauch2012computable,blum2012complexity}.
One of the reason behind this gap is also due to measurement precision,
and some of their computability difference cannot be compensated by
error analysis technique. All these problems for classical computation
could potentially apply to quantum computation. Because we are agnostic
about whether the physical reality is ultimately discrete or continuous,
we tried to develop two different types of quantum theories and computing
models to address these issues. Since the discrete classical computing
model faithfully represents digital computers, we first tried to build
discrete quantum theories and computing by considering quantum theories
and computing over finite fields in Chapter~\ref{chap:DQT=000026DQC}.
Since error analysis technique cannot always compensate the inevitable
problem of measurement precision, we then incorporate the idea of
finite precision measurement into the quantum theory in Chapter~\ref{chap:QIVPM},
and hope it could describe the physical reality and computability
more faithful.

\chapter{Conventional Quantum Theory and Computing}

The part of conventional quantum theory (CQT) used by quantum circuit
model is described by the following:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:-orthonormal-basis}$D$ orthonormal basis vectors for
a Hilbert space of dimension~$D$, 
\item \label{enu:probability-amplitude}the normalized $D$ complex probability
amplitude coefficients describing the contribution of each basis vector, 
\item \label{enu:unitary-matrix}a set of probability-conserving unitary
matrix operators that suffice to describe all required state transformations
of a quantum circuit, 
\item \label{enu:measurement-framework}and a measurement framework. 
\end{enumerate}
In Sec.~\ref{CQC1qubitBloch.sec}, we focus on the discrete geometric
issues raised by the properties \ref{enu:-orthonormal-basis} and
\ref{enu:probability-amplitude} given above for CQT. In Sec.~\ref{sec:fuzzy},
we introduce the important issues of \ref{enu:measurement-framework}
and the foundations of quantum probability space. In Sec.~\ref{sec:Quantum-Circuit-Model},
we address the property~\ref{enu:unitary-matrix} by describing product
and entangled $n$-qubit states and unitary matrices in quantum circuit
model.

@@

\begin{figure}[tb]
\hfill{}\subfloat[]{\centering{}\includegraphics[width=0.48\columnwidth]{hopfS2}}\hfill{}\subfloat[]{\begin{centering}
\includegraphics[width=0.48\columnwidth]{hopfFibers}
\par\end{centering}
}\hfill{} \caption{(a) The two-sphere~$\Sphere{2}$ represented by \Eq{A7.eq}, which
is the irreducible space of one-qubit states, along with a representative
set of points on the sphere. Each single point on the sphere in (a)
corresponding to a circle in (b), and a whole family of circles (the
paths of $\rme^{\rmi\theta}$) on the three-sphere~$\Sphere{3}$
represents the Hopf fibration, Eq.~(\ref{eq:HopfFibration}). Although
$\Sphere{3}$ cannot be directly embedded in $\R^{3}$, three-sphere~$\Sphere{3}$
can be regarded as attaching two three-dimensional ball on two sides
of two-sphere~$\Sphere{2}$. In this way, each circle in $\Sphere{3}$
can be represented as a circle in the three-dimensional ball as shown
in (b). Moreover, points in (a) are color coded corresponding to circles
in (b), e.g., one pole contains the red elliptical circle that would
become an infinite-radius circle by a slightly different way to represent
$\Sphere{3}$ in $\R^{3}$, and the opposite pole corresponds to the
large perfectly round red circle at the equator.}
\label{hopfFibers.fig} 
\end{figure}

>>>>

%% LyX 2.3.1-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,twoside]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{csquotes}
\usepackage{comment}
\usepackage{framed}
\usepackage{mathrsfs}
\usepackage{setspace}
\usepackage{dsfont}
\usepackage{xeCJK}
\setCJKmainfont{SimSun}

@@

\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{prop}{Proposition}[chapter]
\newtheorem{thm}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\newtheorem{case}{Case}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{question}{Question}[chapter]
\newtheorem*{method}{Method}
\newtheorem{example}{Example}[chapter]

@@

\usepackage{unicode-math}
\RequirePackage{ifthen}
\@ifpackageloaded{unicode-math}{}{
 \newcommand{\symbf}{\mathbf}
 \newcommand{\symrm}{\mathrm}
 \newcommand{\symcal}{\mathcal}
 \newcommand{\symbb}{\mathbb}
}
\newcommand{\muB}{\ensuremath{\mu^{\symrm{B}}}}
\newcommand{\muC}{\ensuremath{\mu^{\symrm{C}}}}
\newcommand{\muF}{\ensuremath{\mu^{\symrm{F}}}}
\newcommand{\symrmL}{\ensuremath{\symrm{L}}}
\newcommand{\symrmR}{\ensuremath{\symrm{R}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\symrmL{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\symrmR{#1}}}}
\newcommand{\barmuD}{\ensuremath{\bar{\mu}^{\symrm{D}}}}
\newcommand{\events}{\ensuremath{\symcal{E}}}
\newcommand{\eventsC}{\ensuremath{\events_{\symrm{C}}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\symbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\symbf{#1}}}
\newcommand{\mathReal}{\symbb{R}}
\newcommand{\mathComplex}{\symbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\symbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\symrm{e}}
\newcommand{\rmi}{\symrm{i}}
\newcommand{\rmd}{\symrm{d}}
\newcommand{\CP}[1]{\mathComplex\symbf{P}^{#1}} % Complex projective space
\newcommand{\DCP}[1]{\symbf{D}\mathComplex\symbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\Psibar}{\overline{\Psi}}
\newcommand{\cardp}[2]{#1 \, \slash\!\! \slash \, #2}
\def\round{\mathop{\rm round}\nolimits}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\Hilb}{\symcal{H}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\def\C{{\symbb{C}}}
\newcommand{\ultramodular}{\symcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\nb}{\nolinebreak[3] }

@@

\usepackage{datetime2}
\usepackage{datetime2-calc}
\monthGranted{\DTMmonthname{\month}}
\yearGranted{\the\year}

\usepackage{tabularx}

@@

\usepackage{polyglossia}
\setdefaultlanguage[variant=american]{english}
\setotherlanguage{arabic}
\usepackage[style=numeric-comp,sorting=none,maxbibnames=10,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{prop.bib}
\begin{document}
\title{Discrete Quantum Theories and Computing}
\author{Tai, Yu-Tsung}

\maketitle
%data for Acceptance Page
\committeeMember{Amr A. Sabry, PhD} \committeeMember{Dylan Paul
Thurston, PhD} \committeeMember{Gerardo Ortiz, PhD} \committeeMember{Andrew
J. Hanson, PhD} \committeeMember{Shouhong Wang, PhD} \defensedate{November
29, 2018} \acceptancepage

%data for Copyright Page
\copyrightpage

\begin{dedication}To my parents, Cheng-Tien Tai (戴振沺) and Feng-Ming
Chang (張鳳鳴).\end{dedication}

\begin{acknowledgments*}\todo{Confirm people to be acknowledged
like the way being acknowledged... }I want to thank Prof.~Gerardo
Ortiz, Prof.~Amr Sabry (\textarabic{عمرو صبري}), and Prof.~Andrew
Hanson for publishing three papers \cite{geometry2013,DQT2014,THOS2018}
with me, and enormous other things. These papers become the main part
of my thesis, especially most of Chapter~\ref{chap:DQT=000026DQC}
is based on “Geometry of discrete quantum computing”~\cite{geometry2013}
and “Discrete quantum theories” (DQT)~\cite{DQT2014}; most
of Chapter~\ref{chap:QIVPM} is based on “Quantum Interval-Valued
Probability: Contextuality and the Born Rule” (QIVPM) \cite{THOS2018}.
Together with Prof.~Dylan Thurston and Prof.~Shouhong Wang (汪守宏),
I want to thank you for getting together to understand what I did
and passing my dissertation proposal Spring 2017. I also thank Prof.
Dylan Thurston for serving in my advisory committee and my Tier 3
committee, with other advisory committee members passing my Computer
Science qualifying exam Spring 2016. Especially, my survey of real
computation in the qualifying exam inspired me the comparison of quantum
computer and analog computer in Chapter~\ref{chap:Introduction}.
Chapter~\ref{chap:Introduction} also contains the motivation in
my application for Rethinking Foundations of Physics 2017 workshop
reviewed by Prof. Amr Sabry (\textarabic{عمرو صبري}). I want
to thank Prof. Lawrence Moss for inviting me to present our results
in the interdisciplinary logic seminar and theory seminar in Computer
Science, co-chairing my Tier 3 committee, writing an assessment letter
for me, and drawing my attention on Prof.~Abramsky's paper~\cite{Abramsky2012},
which inspired me to merge quantum probability with discrete quantum
theories. And I thank John Gardiner for inspiring discussion on the
difficulties merging them~\cite{Gardiner2014}. These results are
improved and motivated the transition from DQT to QIVPM in Sec.~\ref{sec:Toward-IVPM}.
I also thank Prof.~Tom Lewis for helping me organize the unpublished
deterministic quantum algorithms in the term paper for SLST-T501 Academic
Writing course as presented in Sec.~\ref{subsec:Discrete-Deutsch-algorithm}.
I want to thank Traci Nagle for going through the literature review
about finite precision measurement and contextuality discussed for
CSCI-Y790 Writing and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.
I also thank Elizabeth “Betsy” Merceron and Kexin “Casey”
Chen (谌可心) for teaching me English in general so that others can understand
me more easily.

\todo{Type a table for acknowledgment so that I would have to write
“I thank” so many times here? }Many people in both department
of Mathematics and Computer Science need to do something more to make
my double-major works smoothly. I need to thank all of them, especially
directors of graduate study in Mathematics, especially Prof. Christopher
Judge and Prof. Matthias Weber; directors of Ph.D. study in Computer
Science, especially Prof. Yuqing Melanie Wu (吴愈青) and Prof. Funda
Ergun.

@@

Finally, thank Jin-Ru Yang (楊謹如) for everything could not be written
here.\end{acknowledgments*}

\begin{preface*}
\begin{itemize}
\item The latest version of this thesis is in GitHub:
\begin{verbatim}
https://git.io/fxbuG
\end{verbatim}
\item Its source code is typeset using \LyX ~\cite{LyX} in:
\begin{verbatim}
https://git.io/fxbuE
\end{verbatim}
\item Any comments can be left in the Issue part of the GitHub Repository:
\begin{verbatim}
https://git.io/fxbug
\end{verbatim}
\end{itemize}
\end{preface*}

\begin{abstract*}Most of the quantum computing models are based on
the continuum of real numbers, despite the fact that classical digital
computers faithfully realize the prediction of discrete computational
models, while analog computers in reality are far weaker than the
theoretical prediction of continuous computational models. Since one
of the important factors of this gap is the measurement precision,
we hope to build a more faithful model by investigating discrete quantum
computing models and incorporating the idea of finite precision measurement
into the quantum theory.

We first replace the continuum of complex numbers by discrete finite
fields in the quantum theory. The simplest theory, defined over unrestricted
finite fields, is so weak that it cannot express Deutsch’s algorithm.
This quantum theory is, however, also so powerful that it can be used
to solve the $\usat$ problem, which is as hard as a general $\NP$-complete
problem. The second framework employs only finite fields with no solution
to $x^{2}+1=0$, and thus permits an elegant complex representation
of the extended field by adjoining $\rmi=\sqrt{-1}$. Because the
states of a discrete $n$-qubit system are in principle enumerable,
we can count the number of states, and determine the proportions of
entangled and unentangled states. Although this improved framework
allows the probabilistic Grover search algorithm in a local region,
we haven't found sensible correspondences between quantum probability
measures and the whole state space in general.

To search a sensible correspondence and to incorporate the idea of
finite precision measurement at the same time, we shift our attention
to consider quantum interval-valued probability measures (IVPMs).
This interval-valued framework not only provides a natural generalization
of both classical IVPMs and conventional quantum probability measures,
but also establishes bounds on the validity of the Kochen-Specker
and Gleason theorems in realistic experimental environments.\end{abstract*}

\tableofcontents{}

@@

When we check the compatibility between quantum mechanics and computer
science, we found one of the biggest difference is that the success
of digital computers is based on discrete computation while quantum
physics is usually based on the continuum of real and complex numbers.
This suggests whether the computational model is discrete or continuous
may be another important factor parallel to \todo{Find a phrase better
than “parallel to”. }whether the model is classical or quantum.
\todo{Add discussion about randomness, like: “Another big difference
is that randomness in quantum computing is inherited from its underlying
quantum physics while randomness in probabilistic Turing machine is
an addition feature acquired by adding random bits to a Turing machine.”
}Although building a realistic quantum computer is an ongoing project~\cite{Shieber2018},
reviewing the challenge posed by classical discrete and continuous
models might provide what we are facing when moving on to the quantum
world.

There was a long history that human believe they can compute with
the continuum of real numbers. If the length of a line segment represents
a real number, Euclid thought he could use an unmarked ruler and a
compass to compute the real numbers about 300 B.C~\cite{MCC2006}.
After logarithm was discovered, a slide rule, also known as a slipstick,
is used to compute multiplication, division, and more complex operations
as a mechanical analog computer~\cite{wiki:SlideRule}. Although
more and more complex analog computers, like integrators \cite{wiki:Integrator,wiki:ChargeAmplifier},
is used, there is still a gap between the computation preformed in
an analog computer and the idealized operations among real numbers.
One problem is that the quality computed by an analog computer might
actually be quantized. For example, charge amplifier is believed to
produce the integrated value of the input current~\cite{wiki:ChargeAmplifier}
although the input charge must actually be an integer multiple of
the elementary charge~\cite{wiki:ElementaryCharge}. Even for the
qualities which doesn't known to be quantized, the observed computational
power of an analog computer might still be weaker than the theoretical
prediction because the precision of an analog computer is limited
generally three or four significant figures~\cite{wiki:AnalogComputer}.
For example, Blum, Cucker, Shub, and Smale modeled their machine to
allow the operations of deciding whether a number is greater or equal
to zero over real numbers and whether a number is exactly zero over
complex numbers \cites{BSS1989}{Ziegler2007}{blum2012complexity}.
Since no physical analog computer could branch in these operations
based on a few significant figures, the theoretical prediction of
Blum-Cucker-Shub-Smale (BCSS) machine cannot fully apply on physical
devices.

Being aware that symbols within a computation should not differ to
an arbitrarily small extent~\cite{Turing_1937}, Turing defined his
machine with only finite number of symbols and configurations. One
of the benefit of discrete computation is that it can be reliably
repeated with exact equivalence. Therefore, any digital computer can
faithfully simulate a Turing machine limited only by the available
memory and time.

\begin{table}
\noindent \centering{}\caption{\label{tab:Comparison}Comparison between the classical computational
models and their physical realizations}
\begin{tabularx}{\linewidth}{ccX}
\toprule 
\addlinespace
 & Model & Physical Realization\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
Continuum & BCSS machine & The precision of an analog computer was limited so that an analog computer may not reproduce the theoretical prediction reliably.\tabularnewline\addlinespace
\addlinespace
\addlinespace
Discrete & Turing machine & A digital computer can be reliably repeated with exact equivalence.\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabularx}
\end{table}
The situations for classical computation is summarized in Table~\ref{tab:Comparison},
and similar situations could potentially be applied to quantum computation.
\todo{Introduction cleaning up until here. }Similar to classical
analog computers manipulating, say, the position of substance, the
most widely used model of quantum computing, the quantum circuit model,
manipulates the probability amplitudes to do the computation. \todo{Play
with some cloud quantum computers to see if I can find more support
evidence. }Because we are agnostic about whether probability amplitude
is quantized or can really be valued in the continuum of complex numbers,
we tried to develop two different types of models: the quantum theories
and computing over finite fields are discrete models, and the quantum
interval-valued probability measures incorporate the idea of measurement
precision into the quantum theory. To distinguish the modified quantum
theories from the usual ones, the usual ones will be called the ‘conventional
quantum theory’ where necessary and it will be reviewed in Chapter~\ref{chap:Conventional-Quantum-Theory}~\footnote{Alternative terminology in the literature includes ‘actual,’ ‘standard,’
and ‘ordinary’ quantum theory.}.

Since conventional probability amplitude is in the field of complex
numbers, and the field structure is essential to compute the probability
from some probability amplitudes, we in Chapter~\ref{chap:DQT=000026DQC}
tried to build discrete quantum theories and computing by considering
them over finite fields.

Since error analysis technique cannot always compensate the inevitable
problem of measurement precision, we then incorporate the idea of
finite precision measurement into the quantum theory in Chapter~\ref{chap:QIVPM},
and hope it could describe the physical reality and computability
more faithful.

\chapter{Conventional Quantum Theory\label{chap:Conventional-Quantum-Theory}}

The part of conventional quantum theory (CQT) used by quantum circuit
model is described by the following:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:-orthonormal-basis}$D$ orthonormal basis vectors for
a Hilbert space of dimension~$D$, 
\item \label{enu:probability-amplitude}the normalized $D$ complex probability
amplitude coefficients describing the contribution of each basis vector, 
\item \label{enu:unitary-matrix}a set of probability-conserving unitary
matrix operators that suffice to describe all required state transformations
of a quantum circuit, 
\item \label{enu:measurement-framework}and a measurement framework. 
\end{enumerate}
In Sec.~\ref{CQC1qubitBloch.sec}, we focus on the discrete geometric
issues raised by the properties \ref{enu:-orthonormal-basis} and
\ref{enu:probability-amplitude} given above for CQT. In Sec.~\ref{sec:fuzzy},
we introduce the important issues of \ref{enu:measurement-framework}
and the foundations of quantum probability space. The observables
and expectation values defined in Sec.~\ref{sec:fuzzy} will be used
to understand the geometry of entangled states in Sec.~\ref{sec:The-geometry-of}.
The property~\ref{enu:unitary-matrix} directly related to quantum
circuit will be introduced later in Chapter~\ref{chap:DQT=000026DQC}
when we dive into quantum algorithms.

@@

\begin{figure}[tb]
\hfill{}\subfloat[]{\centering{}\includegraphics[width=0.48\columnwidth]{hopfS2}}\hfill{}\subfloat[]{\begin{centering}
\includegraphics[width=0.48\columnwidth]{hopfFibers}
\par\end{centering}
}\hfill{} \caption{(a) The two-sphere~$\Sphere{2}$ represented by \Eq{A7.eq}, which
is the irreducible space of one-qubit states, along with a representative
set of points on the sphere. Each single point on the sphere in (a)
corresponding to a circle in (b), and a whole family of circles (the
paths of $\rme^{\rmi\theta}$) on the three-sphere~$\Sphere{3}$
represents the Hopf fibration, Eq.~(\ref{eq:HopfFibration}). Although
$\Sphere{3}$ cannot be directly embedded in $\R^{3}$, three-sphere~$\Sphere{3}$
can be regarded as attaching two three-dimensional ball on two sides
of two-sphere~$\Sphere{2}$. In this way, each circle in $\Sphere{3}$
can be represented as a circle in the three-dimensional ball as shown
in (b). Moreover, points in (a) are color coded corresponding to circles
in (b), e.g., one pole contains the red elliptical circle that would
become an infinite-radius circle by a slightly different way to represent
$\Sphere{3}$ in $\R^{3}$, and the opposite pole corresponds to the
large perfectly round red circle at the equator.}
\label{hopfFibers.fig} 
\end{figure}
The resulting manifold is the two-sphere $\Sphere{2}$ (a 2-manifold)
embedded in $\R^{3}$. If we choose one of many possible coordinate
systems describing $\Sphere{3}$ via \Eq{A5.eq} such as 
\begin{equation}
\left(x_{0},y_{0},x_{1},y_{1}\right)=\left(\cos\left(\theta+\phi\right)\cos\psi,\,\sin\left(\theta+\phi\right)\cos\psi,\,\cos\left(\theta-\phi\right)\sin\psi,\,\sin\left(\theta-\phi\right)\sin\psi\right)\,,\label{A8a.eq}
\end{equation}
where %% $0\leq \theta \leq 2 \pi$,  $-\pi \leq \phi \leq  \pi$, 
$0\leq\psi\leq\frac{\pi}{2}$, with $0\leq\theta+\phi<2\pi$ and $0\leq\theta-\phi<2\pi$,
we see that 
\begin{equation}
\left(X,Y,Z\right)=\left(\cos\left(2\phi\right)\sin\left(2\psi\right),\,\sin\left(2\phi\right)\sin\left(2\psi\right),\,\cos\left(2\psi\right)\right)\,.\label{A8b.eq}
\end{equation}
Thus the one-qubit state is independent of $\theta$, and we can choose
$\theta=\phi$ without loss of generality, reducing the form of the
unique one-qubit states to $\ket{\psi_{1}}=\rme^{2\rmi\phi}\cos\psi\ket{0}+\sin\psi\ket{1}$,
and an irreducible state can be represented as a point on a sphere
called the Bloch sphere, as shown in Fig.~\ref{blochSphere.fig}(a).
